---
title: "Baselines"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Baselines}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message = FALSE}
library(kpe)
```

In addition to `kpe()`, the package provides two baseline estimators
of the personalization effect: `train_eval()` and `papd()`.

## Prerequisites

See `vignette("getting-started", package = "kpe")` for installation and the
full learner/backend table. The runnable comparison at the end uses only base-R
learners (`"linear"`, `"simple"`), so it needs no extra packages; the
`policy_tree`/`random_forest` calls shown for illustration are not evaluated,
but running them would require **`policytree`**, **`ranger`**, and **`glmnet`**.

## `train_eval()`

```{r trev-call, eval = FALSE}
res <- train_eval(X, A, Y, propensity,
                  n_shuffles        = 100,
                  contextual_policy = "policy_tree",
                  best_arm          = "ips",
                  reward_model      = "random_forest",
                  seed              = 0)
```

**Inputs.** Identical to `kpe()`. The `n_folds` argument is accepted
for symmetry and is ignored.

**Output.** An S3 list of class `"kpe"`. `policy_stability` and
`overall_stability` are returned as `NaN` because each shuffle
evaluates a different randomly-chosen held-out half, so per-sample
policy agreement across shuffles is not well-defined.

**Computation.** On each of `n_shuffles` random 50/50 splits of the
data, the reward, contextual-policy, and best-arm learners are
estimated on one half and the AIPW influence function of Eq. S17 is
evaluated on the other half. The Algorithm-1 variance decomposition
is applied to the per-shuffle `(psi_hat, sigma_hat)` pairs using
`n / 2` as the effective sample size for the t-test denominator.

## `papd()`

```{r papd-call, eval = FALSE}
res <- papd(X, A, Y, propensity,
            n_shuffles        = 100,
            contextual_policy = "policy_tree",
            best_arm          = "ips",
            seed              = 0)
```

**Inputs.** Identical to `kpe()`. The `n_folds` and `reward_model`
arguments are accepted for symmetry and are ignored; `papd()` always
uses 3 folds and does not fit a reward model.

**Output.** An S3 list of class `"kpe"`.

**Computation.** Implementation of the Population Average
Prescriptive Difference of Imai and Li (2023), specialized to the
case where one policy class is restricted to providing a single best
arm for all individuals, as described in Li and Brunskill (2026,
Science, 2026). On each of `n_shuffles` random 3-fold partitions, the
contextual policy and best-arm learner are trained on two folds and
the test-fold-centered inverse-propensity statistic

$$\psi^{\text{PAPD}}_i = \frac{\mathbf{1}\{A = \pi(X), \pi_0 \neq \pi(X)\}
                              - \mathbf{1}\{A = \pi_0, \pi_0 \neq \pi(X)\}}
                           {p(A \mid X)}\,(Y_i - \bar Y_{\text{test}})$$

is computed on the held-out fold. The Algorithm-1 variance
decomposition is applied to the per-shuffle `(psi_hat, sigma_hat)`
pairs.

## Side-by-side comparison on a synthetic dataset

```{r compare}
set.seed(1)
n <- 400
X <- matrix(runif(n * 2, -1, 1), ncol = 2)
A <- rbinom(n, 1, 0.5)
Y <- ifelse(X[, 1] >= 0, A, 0.5 * (1 - A)) + rnorm(n, sd = 0.3)

common <- list(
  propensity        = 0.5,
  n_shuffles        = 10,
  contextual_policy = "linear",
  best_arm          = "simple",
  reward_model      = "linear",
  seed              = 0
)
r_kpe  <- do.call(kpe,        c(list(X = X, A = A, Y = Y), common))
r_trev <- do.call(train_eval, c(list(X = X, A = A, Y = Y), common))
r_papd <- do.call(papd,       c(list(X = X, A = A, Y = Y), common))

summary_row <- function(r) {
  ci <- r$confidence_interval
  data.frame(method  = r$method,
             psi     = r$psi,
             ci_lo   = ci[1],
             ci_hi   = ci[2],
             p_value = r$p_value)
}
do.call(rbind, lapply(list(r_kpe, r_trev, r_papd), summary_row))
```
