---
title: "User-supplied learners"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{User-supplied learners}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The `reward_model`, `contextual_policy`, and `best_arm` arguments of
`kpe()`, `train_eval()`, and `papd()` each accept either a character
string naming a built-in learner or a user-supplied object that
implements the required `fit` / `predict` methods.

## Prerequisites

See `vignette("getting-started", package = "kpe")` for installation and the
full learner/backend table. The runnable example below uses only base-R
learners (`"linear"`, `"simple"`) plus a user-supplied reward object, so it
needs no extra packages. The optional `xgboost` snippet near the end requires
the **`xgboost`** package.

## Built-in strings

| slot                | strings                                                                 |
|---------------------|-------------------------------------------------------------------------|
| `reward_model`      | `"random_forest"`, `"linear"`, `"ridge"`, `"lasso"`                    |
| `contextual_policy` | `"policy_tree"`, `"policy_forest"`, `"linear"`                          |
| `best_arm`          | `"ips"`, `"simple"`                                                    |

## User-supplied-object contracts

A user-supplied learner must be an R `list` with `fit` and `predict`
elements matching the following signatures.

```r
# Reward model
list(
  fit     = function(X, y) { ... ; invisible(NULL) },
  predict = function(X) numeric_vector
)
# X is an (n, d + 1) numeric matrix whose last column is the observed
# action. predict returns a length-n numeric vector of predicted
# outcomes.

# Contextual policy
list(
  fit     = function(X, A, Y, n_actions) { ... ; invisible(NULL) },
  predict = function(X) integer_vector
)
# predict returns a length-n integer vector of actions in
# {0, ..., n_actions - 1}.

# Best-arm (non-contextual)
list(
  fit     = function(A, Y, propensities, n_actions) { ... ; invisible(NULL) },
  predict = function() integer_scalar
)
# predict returns a single integer action in {0, ..., n_actions - 1}.
```

The same `list` instance is reused across folds, so `fit` should reset
any state it needs from the training arguments — for example via a
closure that assigns into its parent environment with `<<-`.

## Example: a user-supplied reward model

The learner below records the mean of `y` at fit time and returns it
as the reward prediction at predict time. It has no statistical value;
it is the minimal object that satisfies the reward-model contract.

```{r setup, message = FALSE}
library(kpe)
```

```{r byo-reward}
byo_reward <- local({
  mu <- NULL
  list(
    fit     = function(X, y) { mu <<- mean(y); invisible(NULL) },
    predict = function(X) rep(mu, nrow(as.matrix(X)))
  )
})

set.seed(0)
n <- 300
X <- matrix(runif(n * 2, -1, 1), ncol = 2)
A <- rbinom(n, 1, 0.5)
Y <- ifelse(X[, 1] >= 0, A, 0.5 * (1 - A)) + rnorm(n, sd = 0.3)

fit <- kpe(X, A, Y, propensity = 0.5,
           n_shuffles = 4, n_folds = 6,
           reward_model      = byo_reward,
           contextual_policy = "linear",
           best_arm          = "simple",
           seed = 0)
print(fit)
```

## Example: wrapping an external regressor as a reward model

Any regressor that exposes `fit(X, y)` and `predict(X)` can be adapted
by assigning its fitted state in the closure and calling `predict` at
prediction time. The snippet below wraps `xgboost::xgboost` (not a
hard dependency of `kpe`; load separately if required).

```r
library(xgboost)

byo_xgb <- local({
  model <- NULL
  list(
    fit = function(X, y) {
      model <<- xgboost::xgboost(data = as.matrix(X), label = y,
                                 nrounds = 50, verbose = 0)
      invisible(NULL)
    },
    predict = function(X) {
      as.numeric(stats::predict(model, as.matrix(X)))
    }
  )
})

fit <- kpe(X, A, Y, propensity = 0.5,
           reward_model = byo_xgb, seed = 0)
```

The same construction applies to a user-supplied `contextual_policy`
(providing `fit(X, A, Y, n_actions)` and `predict(X)` returning
integer actions in `{0, ..., n_actions - 1}`) or a user-supplied
`best_arm` (providing `fit(A, Y, propensities, n_actions)` and
`predict()` returning a single integer action).
