---
title: "Getting started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message = FALSE}
library(kpe)
```

This vignette shows the minimal sequence of calls needed to obtain a
personalization-effect estimate from `kpe()`.

> **kpe-r** is the R implementation; the Python version is
> [**kpe-py**](https://github.com/StanfordAI4HI/kpe-py). The package installs and
> loads as `kpe` (`library(kpe)`) — "kpe-r" is the project/repository name.

## Install

Installed from GitHub — not yet on CRAN (planned):

```r
# install.packages("pak")
pak::pak("StanfordAI4HI/kpe-r")
# or: remotes::install_github("StanfordAI4HI/kpe-r")
```

## Learners and the packages they need

`kpe()` (and `train_eval()` / `papd()`) plug in three learner "slots" — a
**reward model**, a **contextual policy**, and a **best-arm rule** — plus a
**propensity model** when `is_rct = FALSE`. Each slot takes a string naming a
built-in backend. The core `kpe` package `Imports` only `stats`, `future`, and
`future.apply`; the forest and penalized-regression backends live in `Suggests`,
so **you must install the R packages for the specific learners your call uses.**

| slot | string value | backing R package (function) | required install |
|------|--------------|------------------------------|------------------|
| `reward_model` | `"random_forest"` | **`ranger`** (`ranger::ranger`) | `install.packages("ranger")` |
| | `"linear"` | base R (`stats::lm`) | — (always available) |
| | `"ridge"`, `"lasso"` | **`glmnet`** | `install.packages("glmnet")` |
| `contextual_policy` | `"policy_tree"` (alias `"dr_econml_2"`) | **`policytree`** (`policytree::policy_tree`); its doubly-robust nuisance scores additionally use **`glmnet`** + **`ranger`** | `install.packages(c("policytree","glmnet","ranger"))` |
| | `"policy_forest"`, `"causal_forest"` | **`grf`** (`grf::causal_forest`; binary treatment only) | `install.packages("grf")` |
| | `"linear"` | base R (per-arm `stats::lm`) | — |
| `best_arm` | `"ips"`/`"best_arm_erm"`, `"simple"`/`"simple_best_arm"` | built into `kpe` | — |
| `propensity_model` (only if `is_rct = FALSE`) | `"logistic"` | **`glmnet`** | `install.packages("glmnet")` |
| | `"random_forest"` | **`ranger`** | `install.packages("ranger")` |

For example, the [JobCorps](jobcorps.html) vignette uses
`reward_model = "random_forest"` and `contextual_policy = "policy_tree"`, so it
needs:

```r
install.packages(c("ranger", "policytree", "glmnet"))
```

> **Which "random_forest"?** In this R package, `reward_model = "random_forest"`
> is the **`ranger`** implementation. That is a *different* random-forest library
> from the sibling Python package `kpe-py`, whose `"random_forest"` is
> scikit-learn's `RandomForestRegressor`. Likewise R's `"policy_tree"`
> (`policytree`) differs from Python's `"policy_tree"` (econml `DRPolicyTree`).
> The two ports agree on `linear`/`lasso`-style learners but are **not** bitwise
> identical on the forest/tree learners.

## Example

```{r example, message = FALSE}
set.seed(0)
n <- 400
X <- matrix(runif(n * 2, -1, 1), ncol = 2)
A <- rbinom(n, 1, 0.5)
Y <- ifelse(X[, 1] >= 0, A, 0.5 * (1 - A)) + rnorm(n, sd = 0.3)

fit <- kpe(X, A, Y, propensity = 0.5,
           n_shuffles        = 30,
           n_folds           = 6,
           contextual_policy = "linear",
           best_arm          = "simple",
           reward_model      = "linear",
           seed              = 0)
print(fit)
```

## Output fields

The returned object is an S3 list of class `"kpe"` with the following
elements:

| field                          | description                                                                                   |
|--------------------------------|-----------------------------------------------------------------------------------------------|
| `psi`                          | Point estimate of the personalization effect on the outcome scale.                            |
| `std_error`                    | Standard error of `psi`.                                                                      |
| `confidence_interval`          | 95% Wald confidence interval, `c(lo, hi)`.                                                    |
| `p_value`                      | One-sided upper-tail t-test p-value.                                                          |
| `psi_per_shuffle`              | Per-shuffle `psi_hat` values.                                                                 |
| `sigma_per_shuffle`            | Per-shuffle `sigma_hat` values.                                                               |
| `policy_stability`             | Fraction of samples for which every shuffle agreed on the contextual-policy action.           |
| `overall_stability`            | Same fraction for the best-arm policy.                                                        |
| `predicted_policy_actions`     | Length-`n` contextual-policy actions from the final shuffle.                                  |
| `predicted_overall_actions`    | Length-`n` best-arm actions from the final shuffle.                                           |
| `method`, `n_samples`, `n_shuffles`, `n_folds` | Run metadata.                                                                                 |

`print`, `summary`, `confint`, and `coef` are defined for the
`"kpe"` class:

```{r methods}
confint(fit)
coef(fit)
```

## Input specification

| argument          | type                          | description                                                                        |
|-------------------|-------------------------------|------------------------------------------------------------------------------------|
| `X`               | numeric matrix or data frame  | `n × d` covariate matrix.                                                          |
| `A`               | integer vector                | Length-`n` observed action in `{0, ..., K-1}`. Factors are coerced.                |
| `Y`               | numeric vector                | Length-`n` observed outcome.                                                       |
| `propensity`      | scalar, vector, or matrix     | `p(A \| X)` for the observed action. Scalars and length-`n` vectors are broadcast. |
| `policy_features` | integer vector, optional      | 1-based column indices of `X` passed to the contextual-policy learner.             |

## Further reading

- [Methods](methods.html) — the algorithm and variance decomposition.
- [Baselines](baselines.html) — the `train_eval()` and `papd()`
  estimators.
- Dataset vignettes: [JobCorps](jobcorps.html), [Joke](joke.html).
- [User-supplied learners](byo.html) — protocols for plugging custom
  `fit` / `predict` objects into each slot.
