---
title: "JobCorps"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{JobCorps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The code assumes that you have created a KPE_DATA_DIR 
env variable which points to the directory you are storing the 
provided data file: jobcorps_cleaned_nonan_full_with_offset.csv. 

This dataset is a semi-synthetic dataset built from the JobCorps 
dataset on job training
(n = 10&nbsp;214; binary treatment; outcome = week-4-year earnings
rate, denominated in dollars per week). 

See details here:
https://datadryad.org/dataset/doi:10.5061/dryad.bg79cnpp7

## References:

* Zhaoqi Li, Emma Brunskill, A statistical test for the benefits of personalizing interventions. Science 393, eaeb9506 (2026). DOI: [10.1126/science.aeb9506](https://doi.org/10.1126/science.aeb9506)
* 2-fold variant detailed [here](https://drive.google.com/file/d/1mmFIdJEQCKhO9Z2flWloWx3ZRQB4P3r4/view)
* Schochet, Peter Z., Burghardt, John, and McConnell, Sheena. Replication data for: Does Job Corps Work? Impact Findings from the National Job Corps Study. Nashville, TN: American Economic Association [publisher], 2008. Ann Arbor, MI: Inter-university Consortium for Political and Social Research [distributor], 2019-10-12. [https://doi.org/10.3886/E113269V1](https://doi.org/10.3886/E113269V1)

## Prerequisites

This vignette assumes `kpe` is installed with the learner backends its call
uses. It calls `reward_model = "random_forest"` and
`contextual_policy = "policy_tree"`, which require the **`ranger`**,
**`policytree`**, and **`glmnet`** packages. See
`vignette("getting-started", package = "kpe")` for installation and the full
learner/backend table.

It also requires the JobCorps CSV and the `KPE_DATA_DIR` environment variable
pointing to the directory that contains it (see above); the code chunks below
are skipped when `KPE_DATA_DIR` is unset.

This dataset can be accessed by cloning the repository and looking in the "vignettes" directory.
git clone https://github.com/StanfordAI4HI/kpe-r.git


## Setup

```{r setup, eval = nzchar(Sys.getenv("KPE_DATA_DIR"))}
library(kpe)

# One-hot every covariate column (mirrors the OneHotEncoder loop in the
# Python JobCorps vignette), reading the same cleaned CSV. Produces:
#   X                : one-hot encoded covariates
#   A                : 0/1 treatment
#   Y                : continuous earnings outcome
#   age_group_index  : 1-based column indices of the AGEGROUP one-hots
df <- read.csv(file.path(Sys.getenv("KPE_DATA_DIR"),
                         "jobcorps_cleaned_nonan_full_with_offset.csv"))
Y <- df$EARNY4
A <- as.integer(df$TREATMNT)

feature_columns <- setdiff(names(df), c("EARNY4", "TREATMNT"))
encoded <- list()
age_group_cols <- character(0)
for (col in feature_columns) {
  lv <- sort(unique(df[[col]]))            # sklearn OneHotEncoder sorts categories
  mm <- outer(df[[col]], lv, `==`) + 0     # full one-hot, no reference level dropped
  colnames(mm) <- paste0(col, "_", lv)
  encoded[[col]] <- mm
  if (col == "AGEGROUP") age_group_cols <- colnames(mm)
}
X <- do.call(cbind, encoded)
X[is.na(X)] <- 0
age_group_index <- match(age_group_cols, colnames(X))   # 1-based indices

pa_x <- prop.table(table(A))[as.character(A)]
```

## Run

```{r run, eval = nzchar(Sys.getenv("KPE_DATA_DIR"))}
result <- kpe(
  X = X, A = A, Y = Y, propensity = pa_x,
  n_shuffles        = 100,
  n_folds           = 6,
  contextual_policy = "policy_tree",     # alias of research-code 'dr_econml_2'
  best_arm          = "best_arm_erm",
  reward_model      = "random_forest",
  policy_features   = age_group_index,
  n_cores           = 5,
  seed              = 0
)
summary(result)
```

## Results


| quantity                       | value                 |
|--------------------------------|-----------------------|
| `psi`                          | `+13.06538`           |
| `std_error`                    | `2.58630`             |
| t-statistic                    | `5.05`                |
| one-sided p-value              | `2.2e-07`             |
| `policy_stability`             | `1.000`               |
| best-arm stability             | `0.810`               |

The confidence interval excludes zero at the 5% level.
`policy_stability = 1.000` records that every shuffle agreed on the
contextual-policy action for every sample: under the default
leave-one-fold-out scheme the exact `policytree` fit on 5/6 of the data
is perfectly consistent across shuffles. 

Note that relative to the results reported in Li and Brunskill 2026
(see their Algorithms 1 and 2), this code base uses a
two-fold shared-nuisance approach in which every nuisance is trained
on 5/6 of the data. This improves the point estimate and the stability,
but the sign  and reject-at-0.05 decision are unchanged.


## Baselines on the same inputs

```{r baselines, eval = nzchar(Sys.getenv("KPE_DATA_DIR"))}
common <- list(
  X = X, A = A, Y = Y, propensity = pa_x,
  n_shuffles        = 100, n_folds = 6,
  contextual_policy = "policy_tree",
  best_arm          = "best_arm_erm",
  reward_model      = "random_forest",
  policy_features   = age_group_index,
  n_cores           = 5, seed = 0
)
r_trev <- do.call(train_eval, common)
r_papd <- do.call(papd,       common)
```

Both baselines return annualized `psi` and `std_error` on the same
scale as `kpe()`.

## Notes

- The JobCorps CSV stores earnings in dollars per week. Figure 3
  multiplies the estimator's output by 52 for presentation; this
  vignette reports the annualized scale for direct comparison with
  the paper.
- `policy_features = age_group_index` restricts the contextual policy
  to the AGEGROUP one-hot columns.
